---
title: "Data Matrices"
author: "Emily Bishop"
date: "`r Sys.Date()`"
output: html_document
link-citations: true
---

```{r setup, echo=FALSE, message = FALSE}
library(here)

load(here("data", "fish.list.Rdata")) 
```

The following matrices were constructed as a result of the decisions described on the home page. 

### The R (environmental) matrix
For this analysis, we're focusing on indices of urbanization that are likely to impact the functional traits of the nearshore community. This matrix will have rows associated with each site and columns for urbanization indices within a buffer of the shoreline. 

One potential approach would be to crop NOAA C-CAP data by USGS HUC-12 (watershed) to get the % developed, % natural, and % agricultural area in the watershed, in addition to % armor of surrounding shorelines. 

### The L (community/species distribution) matrices
For fish, abundances will be pooled across depths within sites. There were a couple of instances where we didn't sample at every shoreline within a site so the abundance matrix needs to account for differences in survey effort. We also didn't sample at every site in every month each year. 

Approach: treat ipas (shorelines) as replicates? calculate average number of each species caught per sampling event (single day)?

```{r load_fish_abund}
head(fish.list[[2]])
```


### The Q (trait) matrices

Fish:
```{r load_fish_traits}
head(fish.list[[1]])
```